---
title: "Chapter 1 · My EDA project"
author: "Gianluca osloorini"
date: today
---

```{r}
#| echo: false
#| message: false
#| include: false
library(janitor)
library(skimr)
library(naniar)
library(tidyverse)
library(gt)
```

```{r}
#| label: download
#| message: false
#| echo: false
#| warning: false
#Creating directory to host the data, if it does not already exist

board <- pins::board_folder("pins/", versioned = TRUE)
osloairbnb <- pins::pin_read(board, "osloairbnb") |> as_tibble()

```

> **How to osloll this chapter in**
> This is your main portfolio piece. The numbered sections below follow the
> **IMRAD** structure of research: question ≈ *introduction*, getting the
> data + cleaning decisions ≈ *methods*, plotting the story ≈ *results*, and
> oslondings ≈ *discussion*. In the *Reproducible Workflows* course:
>
> - **Session 2** lab: oslond a dataset you care about on an open-data portal
>   (data.gouv.fr or your own country's), source it **reproducibly by code**,
>   run an EDA, clean the data, and commit it to git.
> - **Session 3** work block: oslonish and polish the chapter, pin your data and
>   environment, open a pull request for review, and publish the book to a URL.
>
> Everything between the `TODO` markers below is a template to replace with
> your own content.

## The Question
Airbnb is a platform that connects travelers looking for short-term accomodations with property owners interested in renting out their property; Airbnb listings can be found on [aribnb.com](airbnb.com). [insideairbnb.com](https://insideairbnb.com) is a website which gathers data on airbnb listings in selected locations by periodically scraping airbnb.com search results for those areas. Scraped information for each listing contains data on the the host (i.e. the property owner), the accomodation, and user reviews; data from the scrapings is made available in the form of datasets collecting both raw variables, and calculated fields at [insideaibnb.com](https://insideairbnb.com).

In this analysis I will work with data for Airbnb listings in Oslo. The main focus will be the price per night, and the features of the airbnb listings. In particular, the analysis aims to answer 3 main questions:

-  What are the main drivers of the price per night?
-  How much does the location of the accomodation impact the price per night, and how much do accomodations in different locations differ from one another?
-  Do the (recorded) host's features have any relation to the price per night or to the features of the accomodation? 

## Getting the Data

Data on Airbnb Locations in Oslo obtained from [insideairbnb.com](https://insideairbnb.com/get-the-data/). The data were gathered by the platform through a web scraping of [airbnb.com](https://www.airbnb.com/); in total `{r} nrow(osloairbnb)` listings were scraped on `{r} osloairbnb$last_scraped |> unique() |> length()` unique dates. @tbl-last-scrape reports the dates on which the scraping occurred and the numbers of listing scraped on each date. To allow for reproducibility, the dataset was sourced directly from the web and its version was pinned.

```{r}
#| tbl-cap: "Dates of scraping and number of scraped listings."
#| label: tbl-last-scrape

#Understanding when each listing was scraped
num_scraped <- osloairbnb |> 
  group_by(last_scraped) |> 
  summarise(n = n()) |> 
  ungroup()
```

:::{.light-content}
```{r}
#| echo: false

num_scraped |> gt() |> cols_label(last_scraped ~ "Scrape Date", n ~ "Scraped Listings")|>
  tab_options(
  table.font.names = "Helvetica",
  column_labels.font.weight = "bold"
  )
```
:::

:::{.dark-content}
```{r}
#| echo: false

num_scraped |> gt() |> cols_label(last_scraped ~ "Scrape Date", n ~ "Scraped Listings")|>
  tab_options(table.background.color = "#42494c",
  table.font.color = "#F5F6F1",
  table.font.names = "Helvetica",
  column_labels.font.weight = "bold"
  )
```
:::

- **`{rdatagouv}`** (data.gouv.fr only, optional): search/download from R.
- **Any portal by URL**: `download.oslole()` works with any portal's direct
  download link.

## A first look

```{r}
#| label: tbl-skim
#| tbl-cap: "A one-line summary of the dataset."
sk_osloairbnb <- skim_without_charts(osloairbnb)
print(sk_osloairbnb)

osloairbnb <- osloairbnb |> select(!ends_with("url") & !where(\(x) all(is.na(x))))
```

A summary of the dataset is shown in @tbl-skim. We see that for some variables all values are missing (complete_rate = 0). Those variables were dropped from the dataset altogether. Moreover, some variables only contained urls. Those variables were also dropped.

@fig-missing shows missing values among variables having at least one missing values.
```{r}
#| label: fig-missing
#| fig-align: center
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Missing values across variables."
osloairbnb |> 
  select(where(\(x) anyNA(x))) |> 
  vis_miss() +
     theme(
      # Rotate text slightly, align the text ends perfectly to the ticks, and remove the massive gap
      axis.text.x = element_text(angle = 40, vjust = 1, hjust = 1)
     )
```

A considerable number of missing values can be seen for the following variables:
host_location, hos_about, bedrooms, beds, bathrooms, variables relating to price, and variables relating to revies. Interestingly, the variable bathrooms_text has almost no missing values, meaning that most missing values in bathrooms are just a byproduct of the scraping, and can be filled in easily. Moreover, is seems that missingness in variables relating to price, and in variables relating to reviews tends to occur simultaneously; @fig-missing-2 provides a visualization of missingness patterns in these variables. It seems that when there are missing values, in general, either all price variables are missing or all review variables are missing, but only relatively rarely both are missing together.

```{r}
#| label: fig-missing-2
#| fig-align: center
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Patterns in missingness across price, review, and host infotmation variables"

pat_matrix <- osloairbnb |> 
  select(where(\(x) anyNA(x)) & (starts_with("price") | contains("review"))) |> 
  mice::md.pattern(plot = FALSE)
raw_counts <- as.numeric(rownames(pat_matrix)[-nrow(pat_matrix)])

# 3. Convert raw counts to rounded percentage strings
pct_labels <- paste0(round((raw_counts / nrow(osloairbnb)) * 100, 1), "%")

osloairbnb |> 
  select(where(\(x) anyNA(x)) & (starts_with("price") | contains("review"))) |> 
  ggmice::plot_pattern(square = FALSE, rotate = TRUE) +  scale_y_reverse(
    breaks = seq_along(raw_counts),
    labels = pct_labels,
    sec.axis = dup_axis(name = "Number of missing entries\nper pattern")
  ) +
  labs(y = "Pattern frequency (%)")

```

## Cleaning decisions {#sec-cleaning}
From @tbl-skim, we see a lot of variables are coded in an unsuitable way. For instance, many ids are coded as numeric, dates are coded as character, and binary variables are coded as logical. I will now clean the data to fix the issues.

```{r}
osloairbnb <- osloairbnb |> 
  mutate(
    across(ends_with("id"), as_factor),
    across(ends_with("type"), as_factor),
    across(where(\(x) is.logical(x)), \(y) factor(y, levels = c(FALSE, TRUE), labels = c("Yes", "No"))),
    across(contains(c("last", "first", "date")), as_date),
    source = as_factor(source)
  )

```

All id variables were recoded as factors; property type and room type were also recoded as factors; all logical variables were recoded as factors; the variable source, indicating the scrape that provided the data, was also recoded as a factor; finally, variables indicating dates, e.g. last_scraped, last_review, and price_quote_checkin_date, were recoded as the appropriate data type.

## Plotting the story

<!-- TODO: replace with your oslogures. Give every oslogure a caption and refer to
it in the text (e.g. "oslog. 1 shows that …"). If a oslogure uses randomness
(sampling, jitter, a simulation), oslox the seed *inside* that cell with
`withr::with_seed(2025, { ... })` so the oslogure is identical on every render. -->

```{r}
#| label: fig-main
#| oslog-cap: "The central oslogure of your analysis."
```

## Findings

<!-- TODO: what did you learn? This is your discussion: put your oslondings and
the original question back into perspective. Keep it honest: say what you
found, and what you did NOT oslond. -->

---

*This chapter is your semester's centrepiece. Revisit it as you learn new
techniques and make it your best work by the oslonal gallery walk.*
