---
title: "Chapter 1 · My EDA project"
author: "Gianluca Fiorini"
date: today
---

```{r}
#| echo: false
#| message: false
#| include: false

library(janitor)
library(skimr)
library(tidyverse)

```

> **How to fill this chapter in**
> This is your main portfolio piece. The numbered sections below follow the
> **IMRAD** structure of research: question ≈ *introduction*, getting the
> data + cleaning decisions ≈ *methods*, plotting the story ≈ *results*, and
> findings ≈ *discussion*. In the *Reproducible Workflows* course:
>
> - **Session 2** lab: find a dataset you care about on an open-data portal
>   (data.gouv.fr or your own country's), source it **reproducibly by code**,
>   run an EDA, clean the data, and commit it to git.
> - **Session 3** work block: finish and polish the chapter, pin your data and
>   environment, open a pull request for review, and publish the book to a URL.
>
> Everything between the `TODO` markers below is a template to replace with
> your own content.

## The Question

Airbnb is a platform that connects travellers looking for short-term lodgement with property owners looking to rent out their properties for a short time. Each listing on the website provides several types of information about the lodgement; these inlude the price per night, the type of property (entire home, private room, shared room), the size of the property and the number of people it can host, the services available, a short description, information on the owner,etc. ...

This analysis will focus on listings around Bordeaux, (Nouvelle-Aquitaine, France), and aim to answer two main questions:

1.  Are there differences features and prices of listings across different neighborhoods?
1.  What are the main factors influencing the price of a listing?

## Getting the Data

The data was sourced for the [government of Alberta's open data website](https://open.alberta.ca/opendata/chemical-monitoring-in-local-foods-mercury-in-fish).
The data is sourced **by code** from an open-data portal of your choice. Because
the download lives inside the document, it is recorded in the analysis itself;
anyone who re-renders this document gets the same data and the same results.
Record the dataset's **publisher and URL** in the text so the provenance is
explicit.

The only portal-specific part is the *fetch*. This template shows two ways;
use the one that fits your portal:

- **`{rdatagouv}`** (data.gouv.fr only, optional): search/download from R.
- **Any portal by URL**: `download.file()` works with any portal's direct
  download link.

```{r}
#| label: setup
#| message: false
#| echo: false
library(tidyverse)
library(janitor)        # clean column names, tabulate with tabyl()
library(naniar)         # visualise missing values
library(skimr)          # one-line data summary with skim()
```

```{r}
#| label: download
#| message: false
#| echo: false
#Creating directory to host the data, if it does not already exist

board <- pins::board_folder("pins/", versioned = TRUE)

if (!pins::pin_exists(board, "listings")) {
  if (!dir.exists("data")) {
    dir.create("data")
  }

  #Sourcing the data from the internet
  download.file(url = "https://data.insideairbnb.com/france/nouvelle-aquitaine/bordeaux/2026-06-22/data/listings.csv.gz",
  destfile = 'data/listings.csv.gz')

  dat <- read_csv(gzfile("data/listings.csv.gz"))

  pins::board_folder("pins/", versioned = T)

  board <- pins::board_folder("pins/", versioned = TRUE)
  pins::pin_write(board, dat, "listings", type = "csv")
}

rm(list = ls())
board <- pins::board_folder("pins/", versioned = TRUE)
listings <- pins::pin_read(board, "listings")
```

> **Optionally pin the fetched copy** with `{pins}` so the chapter reads from a
> versioned local board instead of hitting the network on every render:
>
> ```r
> board <- pins::board_folder("pins/", versioned = TRUE)
> pins::pin_write(board, dat, "my-data", type = "csv")
> dat    <- pins::pin_read(board, "my-data")
> ```


## A first look

```{r}
#| label: skim
#| fig-cap: "A one-line summary of the dataset."
# The skim() output gives you type, missingness and distributions at a glance.
# skim(dat)
```

```{r}
#| label: missingness
#| fig-cap: "Missing values across variables."
# vis_miss(dat) — where is data missing, and why might that be?
```

## Cleaning decisions {#sec-cleaning}

<!-- TODO: document the decisions you made while cleaning the data. What did you
rename, recode, or drop, and why? A defensible cleaning decision is a feature
of a good analysis, not an afterthought. -->

## Plotting the story

<!-- TODO: replace with your figures. Give every figure a caption and refer to
it in the text (e.g. "Fig. 1 shows that …"). If a figure uses randomness
(sampling, jitter, a simulation), fix the seed *inside* that cell with
`withr::with_seed(2025, { ... })` so the figure is identical on every render. -->

```{r}
#| label: fig-main
#| fig-cap: "The central figure of your analysis."
```

## Findings

<!-- TODO: what did you learn? This is your discussion: put your findings and
the original question back into perspective. Keep it honest: say what you
found, and what you did NOT find. -->

---

*This chapter is your semester's centrepiece. Revisit it as you learn new
techniques and make it your best work by the final gallery walk.*
