---
title: "Chapter 1 · My EDA project"
author: "Gianluca Fiorini"
date: today
---

```{r}
#| echo: false
#| message: false
#| include: false

library(janitor)
library(skimr)
library(naniar)
library(tidyverse)
```

```{r}
#| label: download
#| message: false
#| echo: false
#| warning: false
#Creating directory to host the data, if it does not already exist

board <- pins::board_folder("pins/", versioned = TRUE)

if (!pins::pin_exists(board, 'Star')) {

  #Sourcing the data from the internet
  data("STAR", package = "AER")

  pins::pin_write(board, STAR, "Star", type = "csv")

  rm(list = c('STAR'))
}

Star <- pins::pin_read(board, "Star")


if (!pins::pin_exists(board, 'star')) {
  #The dataset is provided in wide format but it should be used in long format.
  #To that end we can use the code by the authors of the package.

  ## reshape data from wide into long format
  ## 1. variables and their levels
  nam <- c("star", "read", "math", "lunch", "school", "degree", "ladder",
    "experience", "tethnicity", "system", "schoolid")
  lev <- c("k", "1", "2", "3")
  ## 2. reshaping
  star <- reshape(Star, idvar = "id", ids = row.names(Star),
    times = lev, timevar = "grade", direction = "long",
    varying = lapply(nam, function(x) paste(x, lev, sep = "")))
  ## 3. improve variable names and type
  names(star)[5:15] <- nam
  star$id <- factor(star$id)
  star$grade <- factor(star$grade, levels = lev, labels = c("kindergarten", "1st", "2nd", "3rd"))
  rm(nam, lev)

  pins::pin_write(board, star, "star", type = "csv")

  rm(star)
}

star <- pins::pin_read(board,'star')

```

> **How to fill this chapter in**
> This is your main portfolio piece. The numbered sections below follow the
> **IMRAD** structure of research: question ≈ *introduction*, getting the
> data + cleaning decisions ≈ *methods*, plotting the story ≈ *results*, and
> findings ≈ *discussion*. In the *Reproducible Workflows* course:
>
> - **Session 2** lab: find a dataset you care about on an open-data portal
>   (data.gouv.fr or your own country's), source it **reproducibly by code**,
>   run an EDA, clean the data, and commit it to git.
> - **Session 3** work block: finish and polish the chapter, pin your data and
>   environment, open a pull request for review, and publish the book to a URL.
>
> Everything between the `TODO` markers below is a template to replace with
> your own content.

## The Question
The Student-Teacher Achievement Ratio (STAR) study was a randomized controlled trial conducted in Tennessee elementary schools between the years 1985 and 1989. The main aim of the trial was to understand the impact of class size on students' scores in reading and math to guide policymaking. The study involved `{r} star$id |> unique() |> length()` students across `{r} star$schoolid |> unique() |> length()` schools, and concerned grades K-3. In each school participating in the study, children where assigned randomly to one of three possible class types: "small", "regular", and "regular-with-aide". Small classes contained between 13 and 17 students, while regular classes contained between 22 and 25; regular-with-aide classses contained the same number of students as a regular class but were additionally provided with a full-time paid teaching assistant on top of the regular teacher. Not all students involved in the programme were involved in the study for the entire period; the dataset includes students who have spent at least a year in the programme.

 On top of students' class type and test scores, additional variables were also recorded:
 
 - For teachers: gender, ethnicity, highest academic qualification, years of experience, and position on the career ladder.
 - For pupils: gender, ethnicity, year and quarter of birth, and whether they qualified for free lunch.
 - For schools: type ("inner-city", urban", "suburban", "rural").

## Getting the Data
"Stop and frisk" is a practice of the New York City Police Department (NYPD) where officers temporarily detain, question, and search civilians on the streets for weapons or contraband based on reasonable suspicion. The NYPD publishes records on stop and frisk occurrences yearly on [their website](https://www.nyc.gov/site/nypd/stats/reports-analysis/stopfrisk.page).

The practice has been often criticized over the years for being ineffective (in the sense it leads to few arrests), and for disproportionately be used on people 

The data is sourced **by code** from an open-data portal of your choice. Because
the download lives inside the document, it is recorded in the analysis itself;
anyone who re-renders this document gets the same data and the same results.
Record the dataset's **publisher and URL** in the text so the provenance is
explicit.

The only portal-specific part is the *fetch*. This template shows two ways;
use the one that fits your portal:

- **`{rdatagouv}`** (data.gouv.fr only, optional): search/download from R.
- **Any portal by URL**: `download.file()` works with any portal's direct
  download link.

## A first look

```{r}
#| label: skim
#| fig-cap: "A one-line summary of the dataset."
skim(star)
```

```{r}
#| label: missingness
#| fig-cap: "Missing values across variables."
# vis_miss(dat) — where is data missing, and why might that be?
```

## Cleaning decisions {#sec-cleaning}

<!-- TODO: document the decisions you made while cleaning the data. What did you
rename, recode, or drop, and why? A defensible cleaning decision is a feature
of a good analysis, not an afterthought. -->

## Plotting the story

<!-- TODO: replace with your figures. Give every figure a caption and refer to
it in the text (e.g. "Fig. 1 shows that …"). If a figure uses randomness
(sampling, jitter, a simulation), fix the seed *inside* that cell with
`withr::with_seed(2025, { ... })` so the figure is identical on every render. -->

```{r}
#| label: fig-main
#| fig-cap: "The central figure of your analysis."
```

## Findings

<!-- TODO: what did you learn? This is your discussion: put your findings and
the original question back into perspective. Keep it honest: say what you
found, and what you did NOT find. -->

---

*This chapter is your semester's centrepiece. Revisit it as you learn new
techniques and make it your best work by the final gallery walk.*
