---
title: "Chapter 1 · My EDA project"
author: "Gianluca Fiorini"
date: today
---

```{r}
#| echo: false
#| message: false
#| include: false

library(janitor)
library(skimr)
library(naniar)
library(tidyverse)
library(gt)
```

```{r}
#| label: download
#| message: false
#| echo: false
#| warning: false
#Creating directory to host the data, if it does not already exist

board <- pins::board_folder("pins/", versioned = TRUE)

if (!pins::pin_exists(board, 'trentinoairbnb')) {

  if (!dir.exists("data/")){
    dir.create("data/")
  }

  # Define a standard browser User-Agent
  headers <- c("User-Agent" = "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36")

# Download the file using the header
  download.file(
    url = "https://data.insideairbnb.com/italy/trentino-alto-adige-s%C3%BCdtirol/trentino/2026-06-30/data/listings.csv.gz",
    destfile = "data/listings.csv.gz",
    headers = headers,
    mode = 'wb'
  )


  dat <- read_csv(gzfile("data/listings.csv.gz"))

  pins::pin_write(board, dat, "trentinoairbnb", type = "csv")
  
  rm(list = c("dat", "headers"))
}

trentinoairbnb <- pins::pin_read(board, "trentinoairbnb")

```

> **How to fill this chapter in**
> This is your main portfolio piece. The numbered sections below follow the
> **IMRAD** structure of research: question ≈ *introduction*, getting the
> data + cleaning decisions ≈ *methods*, plotting the story ≈ *results*, and
> findings ≈ *discussion*. In the *Reproducible Workflows* course:
>
> - **Session 2** lab: find a dataset you care about on an open-data portal
>   (data.gouv.fr or your own country's), source it **reproducibly by code**,
>   run an EDA, clean the data, and commit it to git.
> - **Session 3** work block: finish and polish the chapter, pin your data and
>   environment, open a pull request for review, and publish the book to a URL.
>
> Everything between the `TODO` markers below is a template to replace with
> your own content.

## The Question
Objectives: understand the main drivers of price and how prices differ across different areas.

## Getting the Data

Data on Airbnb Locations in trentino obtained from [insideairbnb.com](https://insideairbnb.com/get-the-data/). The data were gathered by the platform through a web scraping of [airbnb.com](https://www.airbnb.com/); in total `{r} nrow(trentinoairbnb)` listings were scraped on `{r} trentinoairbnb$last_scraped |> unique() |> length()` unique dates. @tbl-last-scrape the dates on which the scraping occurred and the numbers of listing scraped on each date.

```{r}
#| tbl-cap: "Dates of scraping and number of scraped listings."
#| label: tbl-last-scrape

#Understanding when each listing was scraped
num_scraped <- trentinoairbnb |> 
  group_by(last_scraped) |> 
  summarise(n = n()) |> 
  ungroup()

num_scraped |> gt() |> cols_label(last_scraped ~ "Scrape Date", n ~ "Scraped Listings") |> opt_stylize(style = 5, color = "red")
```

- **`{rdatagouv}`** (data.gouv.fr only, optional): search/download from R.
- **Any portal by URL**: `download.file()` works with any portal's direct
  download link.

## A first look

```{r}
#| label: fig-skim
#| fig-cap: "A one-line summary of the dataset."
skim(trentinoairbnb)
```

```{r}
#| label: missingness
#| fig-cap: "Missing values across variables."
vis_miss(slice_sample(trentinoairbnb,n =2000))
```

## Cleaning decisions {#sec-cleaning}

<!-- TODO: document the decisions you made while cleaning the data. What did you
rename, recode, or drop, and why? A defensible cleaning decision is a feature
of a good analysis, not an afterthought. -->

```{r}

```

## Plotting the story

<!-- TODO: replace with your figures. Give every figure a caption and refer to
it in the text (e.g. "Fig. 1 shows that …"). If a figure uses randomness
(sampling, jitter, a simulation), fix the seed *inside* that cell with
`withr::with_seed(2025, { ... })` so the figure is identical on every render. -->

```{r}
#| label: fig-main
#| fig-cap: "The central figure of your analysis."
```

## Findings

<!-- TODO: what did you learn? This is your discussion: put your findings and
the original question back into perspective. Keep it honest: say what you
found, and what you did NOT find. -->

---

*This chapter is your semester's centrepiece. Revisit it as you learn new
techniques and make it your best work by the final gallery walk.*
