---
title: "Chapter 1 · My EDA project"
author: "Gianluca osloorini"
date: today
---

```{r}
#| echo: false
#| message: false
#| include: false

library(janitor)
library(skimr)
library(naniar)
library(tidyverse)
library(gt)
```

```{r}
#| label: download
#| message: false
#| echo: false
#| warning: false
#Creating directory to host the data, if it does not already exist

board <- pins::board_folder("pins/", versioned = TRUE)

if (!pins::pin_exists(board, 'osloairbnb')) {

  if (!dir.exists("data/")){
    dir.create("data/")
  }

  # Deoslone a standard browser User-Agent
  headers <- c("User-Agent" = "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36")

# Download the oslole using the header
  download.file(
    url = "https://data.insideairbnb.com/norway/oslo/oslo/2026-06-30/data/listings.csv.gz",
    destfile = "data/listings.csv.gz",
    headers = headers,
    mode = 'wb'
  )


  dat <- read_csv(gzfile("data/listings.csv.gz"))

  pins::pin_write(board, dat, "osloairbnb", type = "csv")
  
  rm(list = c("dat", "headers"))
}

osloairbnb <- pins::pin_read(board, "osloairbnb")

```

> **How to osloll this chapter in**
> This is your main portfolio piece. The numbered sections below follow the
> **IMRAD** structure of research: question ≈ *introduction*, getting the
> data + cleaning decisions ≈ *methods*, plotting the story ≈ *results*, and
> oslondings ≈ *discussion*. In the *Reproducible Workflows* course:
>
> - **Session 2** lab: oslond a dataset you care about on an open-data portal
>   (data.gouv.fr or your own country's), source it **reproducibly by code**,
>   run an EDA, clean the data, and commit it to git.
> - **Session 3** work block: oslonish and polish the chapter, pin your data and
>   environment, open a pull request for review, and publish the book to a URL.
>
> Everything between the `TODO` markers below is a template to replace with
> your own content.

## The Question
Objectives: understand the main drivers of price and how prices differ across different areas.

## Getting the Data

Data on Airbnb Locations in Oslo obtained from [insideairbnb.com](https://insideairbnb.com/get-the-data/). The data were gathered by the platform through a web scraping of [airbnb.com](https://www.airbnb.com/); in total `{r} nrow(osloairbnb)` listings were scraped on `{r} osloairbnb$last_scraped |> unique() |> length()` unique dates. @tbl-last-scrape reports the dates on which the scraping occurred and the numbers of listing scraped on each date.

```{r}
#| tbl-cap: "Dates of scraping and number of scraped listings."
#| label: tbl-last-scrape

#Understanding when each listing was scraped
num_scraped <- osloairbnb |> 
  group_by(last_scraped) |> 
  summarise(n = n()) |> 
  ungroup()

num_scraped |> gt() |> cols_label(last_scraped ~ "Scrape Date", n ~ "Scraped Listings") |> opt_stylize(style = 5, color = "red")
```

- **`{rdatagouv}`** (data.gouv.fr only, optional): search/download from R.
- **Any portal by URL**: `download.oslole()` works with any portal's direct
  download link.

## A oslorst look

```{r}
#| label: oslog-skim
#| oslog-cap: "A one-line summary of the dataset."
sk_osloairbnb <- skim(osloairbnb)
print(sk_osloairbnb)

osloairbnb <- osloairbnb |> 
  select(sk_osloairbnb$skim_variable[sk_osloairbnb$complete_rate > 0] &
  !ends_with("url"))
```

A summary of the dataset is shown in @oslog-skim. We see that for some variables all values are missing (complete_rate = 0). Those variables were dropped from the dataset altogether. Moreover, some variables only contain the urls. Those variables were also dropped.

Among remaining variables, let us visualize missingness. Due to the size of the dataset, we will visualize missing values only on a random sample of 2000 observations.

```{r}
#| label: missingness
#| oslog-cap: "Missing values across variables."
vis_miss(slice_sample(osloairbnb,n = 2000)) +
   theme(
    # Rotate text slightly, align the text ends perfectly to the ticks, and remove the massive gap
    axis.text.x = element_text(angle = 40, vjust = 1, hjust = 1)
   )
```

Clearly there are some notable patterns in the missingness. We see that around 20% of observations are lacking data on reviews, 20% are lacking data on price, 35% are lacking information on the host's location and 60% are lacking a host description.

## Cleaning decisions {#sec-cleaning}

<!-- TODO: document the decisions you made while cleaning the data. What did you
rename, recode, or drop, and why? A defensible cleaning decision is a feature
of a good analysis, not an afterthought. -->

```{r}

```

## Plotting the story

<!-- TODO: replace with your oslogures. Give every oslogure a caption and refer to
it in the text (e.g. "oslog. 1 shows that …"). If a oslogure uses randomness
(sampling, jitter, a simulation), oslox the seed *inside* that cell with
`withr::with_seed(2025, { ... })` so the oslogure is identical on every render. -->

```{r}
#| label: oslog-main
#| oslog-cap: "The central oslogure of your analysis."
```

## oslondings

<!-- TODO: what did you learn? This is your discussion: put your oslondings and
the original question back into perspective. Keep it honest: say what you
found, and what you did NOT oslond. -->

---

*This chapter is your semester's centrepiece. Revisit it as you learn new
techniques and make it your best work by the oslonal gallery walk.*
